Dirac neutrino mass generation from Majorana messenger

### 2019-08-06 draft.tex
lc

%Features of the new propossed solution
In general, solutions for light Dirac neutrino masses require a
continuous symmetry to guarantee their Diracness. This symmetry is
usually identified as the $\operatorname{U}(1)_{B-L}$.  Additionally,
ad-hoc discrete symmetries are invoked to forbid tree level Dirac or
Majorana mass terms for the light right-handed
neutrinos~\cite{Roncadelli:1983ty,Han:2018zcn,Wang:2017mcy}. 
%tree-level without adhoc symmetries
However, tree-level Dirac type-I seesaw with proper choices for the
$\operatorname{U}(1)_{B-L}$ charges have been shown to be consistent
without require any extra ad-hoc discrete symmetry~\cite{Ma:2014qra}.
%one-loop 
In recent works, it has been shown that even for one-loop Dirac
neutrino masses, it is possible to have $\operatorname{U}(1)_{B-L}$ as
the only extra symmetry beyond the SM~\cite{Calle:2018ovc,Bonilla:2018ynb,Saad:2019bqf}\footnote{
%extra-loops
  For extensions with only extra scalars, minimal solutions have been found with two and three loops~\cite{Saad:2019bqf}}.
%conexion with dark matter
As a bonus in this case, the new scalars and fermions circulating the
loop can be dark matter candidates with the stability of the lightest
of them guaranteed by the very same continuous symmetry.
% remark
We focus here in solutions for the radiative Dirac type-I seesaw with
Majorana mediators, which have only an extra symmetry responsible for the
Diracness of the light neutrinos, the absence of any tree-level
mass, and the existence of a dark sector constituted by the
particles circulating the loop.

@@

\section{The model}
\label{sec:Model}
%
\begin{table}[t!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ & $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_X$ \\ \hline
    % $H $  & 2  1/2  &  0 \\
    $\eta$  & $\boldsymbol{2}$ & $1/2$  & $1$ \\
    $S$ & $\boldsymbol{1}$ & $0$  & $2$ \\
    $\sigma$ & $\boldsymbol{1}$ & $0$ & $3$ \\
    \hline
    % $Q_{L_{i}}$  & $(2,-1/6)$ & $0$ \\
    % $\overline{u_{R_{i}}}$ & $(1,-2/3)$ & $0$ \\
    % $\overline{d_{R_{i}}}$ & $(1,1/3)$ & $0$ \\
    % \hline
    % $L_i$  & $(2,-1/2)$ & $0$ \\
    % $\overline{e_{R_i}}$ & $(1,1)$ & $0$ \\
    $\nu_{Ri}$ & $\boldsymbol{1}$ & $0$ & $-4$\\
    $\nu_{R3}$ & $\boldsymbol{1}$ & $0$ & $5$\\
    $\psi_{R\alpha}$  & $\boldsymbol{1}$ & 0 & $1$ \\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges. All the SM fields are neutral under the new $\operatorname{U}(1)_X$ gauge symmetry. }
    \label{tab:partcont}
\end{table}
%
We extend the standard model with a spontaneously broken Abelian gauge symmetry  which guarantees the total lepton number $L$ conservation. Only the new particles including the light right handed neutrinos are charged under this symmetry. We choose the new particle set such that the following dimension six operator is genuinely realized
\begin{align}
  \label{eq:ld6}
  \mathcal{L}_{6D}=\frac{1}{\Lambda^2} \overline{L_i} H \nu_R S^2\,,
\end{align}
where $S$ is the singlet scalar field which can spontanouely break the $\operatorname{U}(1)_X$ symmetry.

With the aim to illustrate the one-loop Dirac neutrino mass generation
we consider the particle content shown in Table \ref{tab:partcont} as
a possible the realization of the effective operator in
ec.~\eqref{eq:ld6}. Specifically, we introduce three scalar fields
($\eta, \sigma$ and $S$), where only $S$ develops a nonzero vacuum
expectation value (VEV), and two sets of three singlet fermions,
$\nu_{Ri}$ and $\nu_{R3}$, and $\psi_{R\alpha}$, where $i=1,2$ and
$\alpha=1,2,3$.  The model also includes an additional $\operatorname{U}(1)_X$ gauge
symmetry which is used to forbid the Dirac and Majorana neutrino mass
terms at tree level.
%Additionally, we add three right neutrinos, one scalar doublet, two scalar singlets and three Majorana fermion singlet under $\operatorname{SU}(2)_L$.
The $\operatorname{U}(1)_X$ charges for the new particles are defined by the anomaly cancellation conditions and the gauge invariance in Yukawa and scalar interactions.
%In this model the neutrinos are Dirac fermions. Particle content and their respective charges are show in Tab.~\ref{tab:partcont}.

@@

A not-so-minimal model can be obtained as the one-loop realization of
the effective 5-dimensional operator for Dirac neutrino masses~\cite{}
\begin{align}
  \label{eq:ld5}
  \mathcal{L}_{5D}=\frac{1}{\Lambda} \overline{L_i} H \nu_R S\,,
\end{align}
as displayed in fig.~\ref{fig:ld5}. In this case we use the usual $\operatorname{U}(1)_{B-L}$ for the lepton doublet, so that an extra set of quiral fermion needs to be added in a vector-like way to cancel out the anomalies. The assignment is displayed in Table~\ref{tab:partcont2}

@@

The two additional left-handed fields $\xi_{L\alpha}$,  with a $\operatorname{U}(1)_{B-L}$ charge of $r=3/2$,  cancel out the anomalies. Moreover, we have both Dirac mass terms $\overline{\xi_{L \alpha}}\psi_{R \beta}$ and $ \overline{\xi_{L\alpha}^c }\xi_{L \beta} \langle S\rangle$ and the full set of Majorana fields are massive and heavy. The lighter of them a would be dark matter particle.

@@

The existence of three new right neutrinos in nature contribute to the total energy density of the universe. These new contributions can modify the cosmological observables in the cosmic microwave background (CMB). These new neutrinos contribute to the existence of dark radiation, which affects the expansion processes of the universe and the number of relativistic degrees of freedom for neutrinos.

In the case of our model, the new gauge symmetry introducesa  new gauge boson that mediates interaction of the right neutrinos and the SM. This allows the right neutrinos to thermally modify the relativistic degrees of freedom with contributions given by~\cite{Anchordoqui:2012qu,Anchordoqui:2011nh}
%
\begin{align*}
    \Delta N_{\text{eff}} = N_{\text{eff}} - N^{SM}_{\text{eff}} = N_{\nu_R} \left( \frac{T_{\nu_{R}}}{T_{\nu_{L}}} \right)^{4} = N_{\nu_R} \left( \frac{g(T^{\nu_{L}}_{\text{dec}})}{g(T^{\nu_{R}}_{\text{dec}})} \right)^{4}
\end{align*}
%
where $ g(T) $ is the number of relativistic degrees of freedom at a temperature $T$ in the SM~\cite{Aghanim:2018eyx}. The decoupling temperature of the left neutrinos $\nu_L$ is $ T^{\nu_{L}}_{\text{dec}} \approx 2.3 $ MeV~\cite{Enqvist:1991gx}, when $g(T^{\nu_{L}}_{ \text{dec}}) = 43/4 $ corresponding to the three $\nu_{L}$, $e^{\pm} $ and the photon~\cite{Kolb:1990vq}. The rate of interaction of the right neutrinos with the SM is mediated by the gauge boson $Z^{\prime} $ and is given by~\cite{SolagurenBeascoa:2012cz}
%
\begin{align*}
    \Gamma_{\nu_R} (T) &= n_{\nu_R}(T) \langle \sigma(\overline{\nu_{R}} \nu_{R} \to \overline{f} f) \upsilon \rangle \,, \\
    &= \frac{g^{2}_{\nu_R}}{n_{\nu_R}(T)} \int \frac{d^{3} p}{(2 \pi)^{3}} f_{\nu_R}(p) \int \frac{d^{3} k}{(2 \pi)^{3}} f_{\nu_R}(k) \sigma_{f}(s) \upsilon\,,
\end{align*}
%
where $f_{\nu_R}(k)=1/(e^{k/T}+1)$ is the Fermi-Dirac distributión, $g_{\nu_R} = 2$ is spin number, $\upsilon = 1-\cos{\theta}$ is the Moller velocity, $s = 2 p k (1-\cos{\theta})$, where $p$ and $k$ are the momenta of the particle and $\theta$, the angle between them and the number density of right-handed neutrinos is given by
\begin{align*}
n_{\nu_R}(T) = g_{\nu_R} \int \frac{d^{3} k}{(2 \pi)^{3}} f_{\nu_R}(k)\,.
\end{align*}

The cross section is given in~\cite{Barger:2003zh}, in the case with heavy mediator $T^{\nu_R}_{\text{dec}} \ll M_{Z^{\prime}}$, we can to take the limit when $s \ll M_{Z^{\prime}}$ and neglecting the fermion masse, we have that the cross section is
%
\begin{align*}
    \sigma_{f}(s) \approx N^{C}_{f} \frac{s}{12 \pi} \left( \frac{g^{\prime}}{M_{Z^{\prime}}} \right)^{4} Q^{2}_{\nu_R} (Q^{2}_{f_L} + Q^{2}_{f_R})\,,
\end{align*}
%
where $N^{C}_{f}$ is the color multiplicity of the fermion, i.e. $N^{C}_{q(\ell)} = 3(1)$, $Q_{f}$ is the charge of the fermion under the new simmetry, $g^{\prime}$ and $M_{Z^{\prime}}$ are the gauge coupling and mass for the new gauge boson. Solving the integrals with this last cross section we obtain that the interaction rate is given by
%
\begin{align*}
    \Gamma_{\nu_{R}}(T) = \frac{49 \pi^{5} T^{5}}{97200 \zeta(3)} \left( \frac{g^{\prime}}{M_{Z^{\prime}}} \right)^{4} \sum_{f} N^{C}_{f} Q^{2}_{f}\,,
\end{align*}
%
%
\begin{figure}
\centering
\includegraphics[scale=0.8]{DeltaNeff.pdf}
\caption{In the upper panel is the decoupling temperature of the right neutrino $\nu_{R}$ in function of $M_{z^{\prime}}/g^{\prime}$. In the lower panel is the additional contribution to the number of relativistic degrees of freedom ($\Delta N_{\text{eff}}$) in function of $M_{z^{\prime}}/g^{\prime}$. The gray shaded region is excluded by the measurements in the Planck CMB~\cite{Aghanim:2018eyx}. The region shaded in violet shows the bound from LEP  LEP~\cite{Alioli:2017nzr}.}
\label{fig:Neff}
\end{figure}
%
where the sum is performed over all SM fermion that are in thermal equilibrium with the plasma at temperature T. To estimate the contribution to relativistic degrees of freedom for neutrinos, it is necessary to calculate the decoupling temperature of the right neutrinos ($T^{\nu_R}_{\text{dec}}$). The latter occurs when the interaction rate of neutrinos with the SM drop below the rate of expansion of the universe, i.e.
%
\begin{align*}
\Gamma(T^{\nu_R}_{\text{dec}}) = H(T^{\nu_R}_{\text{dec}})\,, 
\end{align*}
%
where the Hubble expansion parameter is
%
\begin{align*}
    H(T) = \sqrt{\frac{8 \pi G_{N} \rho(T)}{3}} = \sqrt{\frac{4 \pi^{3} G_{N}}{45} \left( g(T) + \frac{21}{4} \right)} T^{2}\,,
\end{align*}
%
where $21/4$ correspond to the contribution of the right neutrinos to relativistic degrees of freedom. By finding the decoupling temperature of the right neutrinos $T^{\nu_R}_{\text{dec}} $ by a root finding method, it is possible to calculate $\Delta N_{\text{eff}}$.

The results of $T^{\nu_R}_{\text{dec}}$ and $\Delta N_{\text{eff}}$ as a function of $M_{Z^{\prime}}/g^{\prime}$ are exhibited in fig.~\ref{fig:Neff}. We can see that for small decoupling temperatures, the variation in the number of relativistic degrees of freedom of the new species is large and vice versa. The bound that come from Planck~\cite{ghanim:2018eyx}

@@

is stronger than the bound of LEP~\cite{Alioli:2017nzr}. For our model we find that $M_{Z^{\prime}}/g^{\prime} \gtrsim 60$TeV in order to satisfy the conditions of $N_{eff}$.

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "draft"
%%% ispell-local-dictionary: "american"
%%% End: 
The cross section is given in [], in the case with heavy mediator $T^{\nu_R}_{\text{dec}} \ll M_{Z^{\prime}}$, we can to take the limit when $s \ll M_{Z^{\prime}}$ and neglecting the fermion masses, we have that the cross section is

>>>>

%Features of the new propossed solution
In general, solutions for light Dirac neutrino masses require a
continuous symmetry to guarantee their Diracness. This symmetry is
usually identified as the $\operatorname{U}(1)_{B-L}$.  Additionally,
ad-hoc discrete symmetries are invoked to forbid tree level Dirac or
Majorana mass terms for the light right-handed
neutrinos~\cite{Roncadelli:1983ty,Han:2018zcn,Wang:2017mcy}. 
%tree-level without adhoc symmetries
However, tree-level Dirac type-I seesaw with proper choices for the
$\operatorname{U}(1)_{B-L}$ charges have been shown to be consistent
without require any extra ad-hoc discrete symmetries~\cite{Ma:2014qra}.
%one-loop 
In recent works, it has been shown that even for one-loop Dirac
neutrino masses, it is possible to have $\operatorname{U}(1)_{B-L}$ as
the only extra symmetry beyond the SM~\cite{Calle:2018ovc,Bonilla:2018ynb,Saad:2019bqf}\footnote{
%extra-loops
  For extensions with only extra scalars, minimal solutions have been found with two and three loops~\cite{Saad:2019bqf}}.
%conexion with dark matter
As a bonus in this case, the new scalars and fermions circulating the
loop can be dark matter candidates with the stability of the lightest
of them guaranteed by the very same continuous symmetry.
% remark
We focus here in solutions for the radiative Dirac type-I seesaw with
Majorana mediators, which have only an extra symmetry responsible for the
Diracness of the light neutrinos, the absence of any tree-level
mass, and the existence of a dark sector constituted by the
particles circulating the loop.

@@

\section{The model}
\label{sec:Model}
%
\begin{table}[t!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ & $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_X$ \\ \hline
    % $H $  & 2  1/2  &  0 \\
    $\eta$  & $\boldsymbol{2}$ & $1/2$  & $1$ \\
    $S$ & $\boldsymbol{1}$ & $0$  & $2$ \\
    $\sigma$ & $\boldsymbol{1}$ & $0$ & $3$ \\
    \hline
    % $Q_{L_{i}}$  & $(2,-1/6)$ & $0$ \\
    % $\overline{u_{R_{i}}}$ & $(1,-2/3)$ & $0$ \\
    % $\overline{d_{R_{i}}}$ & $(1,1/3)$ & $0$ \\
    % \hline
    % $L_i$  & $(2,-1/2)$ & $0$ \\
    % $\overline{e_{R_i}}$ & $(1,1)$ & $0$ \\
    $\nu_{Ri}$ & $\boldsymbol{1}$ & $0$ & $-4$\\
    $\nu_{R3}$ & $\boldsymbol{1}$ & $0$ & $5$\\
    $\psi_{R\alpha}$  & $\boldsymbol{1}$ & 0 & $1$ \\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges. All the SM fields are neutral under the new $\operatorname{U}(1)_X$ gauge symmetry. }
    \label{tab:partcont}
\end{table}
%
We extend the standard model with a spontaneously broken Abelian gauge symmetry  which guarantees the total lepton number ($L$) conservation. Only the new particles including the light right handed neutrinos are charged under this symmetry. We choose the new particle set such that the following dimension six operator is genuinely realized
\begin{align}
  \label{eq:ld6}
  \mathcal{L}_{6D}=\frac{1}{\Lambda^2} \overline{L_i} H \nu_R S^2\,,
\end{align}
where $S$ is the singlet scalar field which can spontanouely break the $\operatorname{U}(1)_X$ symmetry.

With the aim to illustrate the one-loop Dirac neutrino mass generation
we consider the particle content shown in Table \ref{tab:partcont}. as
a possible realization of the effective operator in
ec.~\eqref{eq:ld6}. Specifically, we introduce three scalar fields
$\eta, \sigma$ and $S$, where only $S$ develops a nonzero vacuum
expectation value (VEV),  two sets of three singlet fermions,
$\nu_{Ri}$ and $\nu_{R3}$, and a set a heavy Majorana fermions, $\psi_{R\alpha}$, where $i=1,2$ and
$\alpha=1,2,3$.  The model also includes an additional $\operatorname{U}(1)_X$ gauge
symmetry which is used to forbid the Dirac and Majorana neutrino mass
terms at tree level.
%Additionally, we add three right neutrinos, one scalar doublet, two scalar singlets and three Majorana fermion singlet under $\operatorname{SU}(2)_L$.
The $\operatorname{U}(1)_X$ charges for the new particles are defined by the anomaly cancellation conditions and the gauge invariance in Yukawa and scalar interactions.
%In this model the neutrinos are Dirac fermions. Particle content and their respective charges are show in Tab.~\ref{tab:partcont}.

@@

A not-so-minimal model can be obtained as the one-loop realization of
the effective 5-dimensional operator for Dirac neutrino masses~\cite{Gu:2006dc}
\begin{align}
  \label{eq:ld5}
  \mathcal{L}_{5D}=\frac{1}{\Lambda} \overline{L_i} H \nu_R S\,,
\end{align}
as displayed in fig.~\ref{fig:ld5}.
In this case we can use the standard $\operatorname{U}(1)_{B-L}$
charge of $-1$ for the lepton doublets.
It is well established that the solution for anomaly cancellation
conditions with three quiral fields with exotic
$\operatorname{U}(1)_{B-L}$ charges: $-4$,~$-4$,~$5$,  when assigned to the
three light right-handed neutrinos can forbid three level Dirac and
Majorana masses~\cite{Calle:2018ovc}.
The presence of a set of heavy Majorana mediators to realize the
dimension-5 operator at one-loop with the diagram in
fig.~\ref{fig:ld5}., would spoil the anomaly cancellation conditions.
However, we can further add an extra set of quiral fermion in a
vector-like way, such that the full set heavy fermions do not affect
the anomaly cancellation. A such assignment is displayed in
Table~\ref{tab:partcont2}.
The fields $\xi_{L \alpha}$ guarantee the anomaly cancellation and
their Majorana masses are obtained in the same way than the ones for
$\psi_{R\alpha}$.

@@

The two additional left-handed fields $\xi_{L\alpha}$,  with a $\operatorname{U}(1)_{B-L}$ charge of $r=3/2$,  cancel out the anomalies. Moreover, we have both Dirac mass terms $\overline{\xi_{L \alpha}}\psi_{R \beta}$ and $ \overline{\xi_{L\alpha}^c }\xi_{L \beta} \langle S\rangle$ and the full set of Majorana fields are massive and heavy. The lighter of them a would be dark matter candidate.

@@

The coupling of $Z'$ with the lepton doublets opens up the possibility
of thermalization of the right handed neutrinos in the primordial
plasma.
In fact, the existence of three new right neutrinos in nature
contribute to the total energy density of the universe. These new
contributions can modify the cosmological observables in the cosmic
microwave background (CMB). These new neutrinos contribute to the
existence of dark radiation, which affects the expansion processes of
the universe and the number of relativistic degrees of freedom for
neutrinos.

In the case of our model, the new gauge symmetry introduces  new gauge boson that mediates interaction of the right neutrinos and the SM. This allows the right neutrinos to thermally modify the relativistic degrees of freedom with contributions given by~\cite{Anchordoqui:2012qu,Anchordoqui:2011nh}
%
\begin{align*}
    \Delta N_{\text{eff}} = N_{\text{eff}} - N^{\text{SM}}_{\text{eff}} = N_{\nu_R} \left( \frac{T_{\nu_{R}}}{T_{\nu_{L}}} \right)^{4} = N_{\nu_R} \left( \frac{g(T^{\nu_{L}}_{\text{dec}})}{g(T^{\nu_{R}}_{\text{dec}})} \right)^{4}
\end{align*}
%
where $ g(T) $ is the number of relativistic degrees of freedom at a temperature $T$ in the SM~\cite{Aghanim:2018eyx}. The decoupling temperature of the left neutrinos $\nu_L$ is $ T^{\nu_{L}}_{\text{dec}} \approx 2.3 $ MeV~\cite{Enqvist:1991gx}, when $g(T^{\nu_{L}}_{ \text{dec}}) = 43/4 $ corresponding to the three $\nu_{L}$, $e^{\pm} $ and the photon~\cite{Kolb:1990vq}. The rate of interaction of the right neutrinos with the SM is mediated by the gauge boson $Z^{\prime} $ and is given by~\cite{SolagurenBeascoa:2012cz}
%
\begin{align*}
    \Gamma_{\nu_R} (T) &= n_{\nu_R}(T) \langle \sigma(\overline{\nu_{R}} \nu_{R} \to \overline{f} f) \upsilon \rangle \,, \\
    &= \frac{g^{2}_{\nu_R}}{n_{\nu_R}(T)} \int \frac{d^{3} p}{(2 \pi)^{3}} f_{\nu_R}(p) \int \frac{d^{3} k}{(2 \pi)^{3}} f_{\nu_R}(k) \sigma_{f}(s) \upsilon\,,
\end{align*}
%
where $f_{\nu_R}(k)=1/(e^{k/T}+1)$ is the Fermi-Dirac distribution, $g_{\nu_R} = 2$ is the spin number, $\upsilon = 1-\cos{\theta}$ is the Moller velocity, $s = 2 p k (1-\cos{\theta})$, where $p$ and $k$ are the momenta of the particle and $\theta$, the angle between them, and the number density of right-handed neutrinos is given by
\begin{align*}
n_{\nu_R}(T) = g_{\nu_R} \int \frac{d^{3} k}{(2 \pi)^{3}} f_{\nu_R}(k)\,.
\end{align*}

The cross section is given in~\cite{Barger:2003zh}, in the case with heavy mediator $T^{\nu_R}_{\text{dec}} \ll M_{Z^{\prime}}$, we can to take the limit when $s \ll M_{Z^{\prime}}$ and neglecting the fermion masses, we have that the cross section is
%
\begin{align*}
    \sigma_{f}(s) \approx N^{C}_{f} \frac{s}{12 \pi} \left( \frac{g^{\prime}}{M_{Z^{\prime}}} \right)^{4} Q^{2}_{\nu_R} (Q^{2}_{f_L} + Q^{2}_{f_R})\,,
\end{align*}
%
where $N^{C}_{f}$ is the color multiplicity of the fermion, i.e. $N^{C}_{q(\ell)} = 3(1)$, $Q_{f}$ is the charge of the fermion under the new symmetry, $g^{\prime}$ and $M_{Z^{\prime}}$ are the gauge coupling and mass for the new gauge boson. Solving the integrals with this last cross section we obtain that the interaction rate is given by
%
\begin{align*}
    \Gamma_{\nu_{R}}(T) = \frac{49 \pi^{5} T^{5}}{97200 \zeta(3)} \left( \frac{g^{\prime}}{M_{Z^{\prime}}} \right)^{4} \sum_{f} N^{C}_{f} Q^{2}_{f}\,,
\end{align*}
%
where the sum is performed over all SM fermion that are in thermal equilibrium with the plasma at temperature $T$. To estimate the contribution to the relativistic degrees of freedom for neutrinos, it is necessary to calculate the decoupling temperature of the right neutrinos ($T^{\nu_R}_{\text{dec}}$). The latter occurs when the interaction rate of neutrinos with the SM drop below the rate of expansion of the universe, i.e.
%
\begin{align*}
\Gamma(T^{\nu_R}_{\text{dec}}) = H(T^{\nu_R}_{\text{dec}})\,, 
\end{align*}
%
where the Hubble expansion parameter is
%
\begin{align*}
    H(T) = \sqrt{\frac{8 \pi G_{N} \rho(T)}{3}} = \sqrt{\frac{4 \pi^{3} G_{N}}{45} \left( g(T) + \frac{21}{4} \right)} T^{2}\,,
\end{align*}
%
and the $21/4$ correspond to the contribution of the right neutrinos to relativistic degrees of freedom. By finding the decoupling temperature of the right neutrinos $T^{\nu_R}_{\text{dec}} $ by a root finding method, it is possible to calculate $\Delta N_{\text{eff}}$.

The results of $T^{\nu_R}_{\text{dec}}$ and $\Delta N_{\text{eff}}$ as a function of $M_{Z^{\prime}}/g^{\prime}$ are exhibited in fig.~\ref{fig:Neff}. We can see that for small decoupling temperatures, the variation in the number of relativistic degrees of freedom of the new species is large and vice versa. The bound that come from Planck~\cite{ghanim:2018eyx} is stronger than the bound of LEP~\cite{Alioli:2017nzr}. For our model we find that $M_{Z^{\prime}}/g^{\prime} \gtrsim 60$TeV in order to satisfy the conditions of $N_{eff}$.

%
\begin{figure}
\centering
\includegraphics[scale=0.8]{DeltaNeff.pdf}
\caption{In the upper panel is the decoupling temperature of the right neutrino $\nu_{R}$ in function of $M_{z^{\prime}}/g^{\prime}$. In the lower panel is the additional contribution to the number of relativistic degrees of freedom ($\Delta N_{\text{eff}}$) in function of $M_{z^{\prime}}/g^{\prime}$. The gray shaded region is excluded by the measurements in the Planck CMB~\cite{Aghanim:2018eyx}. The region shaded in violet shows the bound from LEP  LEP~\cite{Alioli:2017nzr}.}
\label{fig:Neff}
\end{figure}
%

@@

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "draft"
%%% ispell-local-dictionary: "american"
%%% End:
